A Road Segment Prioritization Approach for Cycling Infrastructure

### 2022-08-29 README.Rmd
intro: manchester context

---
title: A Road Segment Prioritization Approach for Cycling Infrastructure
#date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
 # bookdown::github_document2:
   bookdown::pdf_document2:
 # bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options:
  markdown:
    wrap: 72
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: elsevier-harvard.csl
always_allow_html: yes
---

@@

The funding on its own is therefore no guarantee of a change in
commuting across the country; it must be used to design adequate cycling
infrastructure that is based on motivators and deterrents to cycling.

@@

## Network-Level Approaches

@@

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of *connected
components*. @natera2019data for example,
study the existing cycling network in terms of its disconnected
components. They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor. By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
The concept of connected components is
also at the core of the methodology proposed by @olmos2020data. After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

@@

```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="100%", fig.show='hold', fig.align="center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling_potential_current_all.png")
```

@@

Consideration of these limitations suggest future directions of research.
New datasets and continuously evolving computational (hardware and software)
capabilities should enable the data related issues to be overcome as open
datasets, and our ability to process them, improve.
Plugging-in alternative uptake models could help address the relatively narrow
definition of 'potential' used in this paper, to consider broader social factors.
A promising area of future research in this direction could be 
to integrate a wider range of modes, including e-bikes and e-scooters, into the
analysis. Currently scope for progress in this direction is
partially restricted by the lack of data on the
proliferation of these modes, raising the point that surveys could
replace the broad category of `Bicycle' a range of small modes such as
'pedal cycles' (including tricycles and recumbent cycles), e-bikes
and other forms of micro-mobility.

Despite the rapid growth of these alternative modes, there is little doubt that
the humble bicycle is a key ingredient in the ongoing shift towards active transport.
Recognition of the importance of this shift for improved the health and well-being
the citizens, and the wider challenge to decarbonise the global economy,
has grown with pressures on health and public transport systems during the pandemic.
The success of policies to accelerate cycling uptake depends on a range of
factors including, vitally, the design of the
cycling network for potential cyclists.
The approach presented in this paper provides a strong evidence base,
that considers both cycling potential and social equity,
for designing joined-up and cost-effective strategic cycle networks.
<!-- In tandem with policies to reduce excessive reliance on private cars, evidence-based investment in strategic cycle networks is key to. -->

>>>>

---
title: A Road Segment Prioritization Approach for Cycling Infrastructure
author:
  - Hussein Mahfouz:
      email: h.mahfouz.19@ucl.ac.uk
      institute: casa
      correspondence: true
  - name: Elsa Arcaute
    institute: casa
  - name: Robin Lovelace
    institute: its
institute:
  - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
  - its: Institute for Transport Studies, University of Leeds, Leeds, UK

geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
 # bookdown::github_document2:
   bookdown::pdf_document2:
 # bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
      # https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
      pandoc_args:
      - '--lua-filter=extras/scholarly-metadata.lua'
      - '--lua-filter=extras/author-info-blocks.lua'
      keep_tex:  true
editor_options:
  markdown:
    wrap: 72
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: extras/elsevier-harvard.csl
always_allow_html: yes
---

@@

The funding on its own is therefore no guarantee of a change in
commuting across the country; it must be used to design adequate cycling
infrastructure that is based on motivators and deterrents to cycling. In this paper we 
propose an approach for planning a connected cycling network that is meant to 
be applicable to any city, and we use the city of Manchester as a case study. Greater 
Manchester aims to become the first region to reach a carbon neutral transport network; 
Its local authorities have developed plans for a 1800 mile walking and cycling network 
(Bee Network) that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
have already been put in place for around a quarter of this network, the majority is yet to be built.

# Literature Review

@@

## Network-Level Approaches to Bicycle Infrastructure Planning

@@

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of *connected
components*. @orozco2020data for example,
study the existing cycling network in terms of its disconnected
components. They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor. By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
The concept of connected components is
also at the core of the methodology proposed by @olmos2020data. After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

@@

```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="90%", fig.show='hold', fig.align="center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling_potential_current_all.png")
```

@@

Consideration of these limitations suggest future directions of research.
New datasets and continuously evolving computational (hardware and software)
capabilities should enable the data related issues to be overcome as open
datasets, and our ability to process them, improve.
Plugging-in alternative uptake models could help address the relatively narrow
definition of 'potential' used in this paper, to consider broader social factors.
A promising area of future research in this direction could be 
to integrate a wider range of modes, including e-bikes and e-scooters, into the
analysis. Currently scope for progress in this direction is
partially restricted by the lack of data on the
proliferation of these modes, raising the point that surveys could
replace the broad category of `Bicycle' with a range of small modes such as
'pedal cycles' (including tricycles and recumbent cycles), e-bikes
and other forms of micro-mobility.

Despite the rapid growth of these alternative modes, there is little doubt that
the humble bicycle is a key ingredient in the ongoing shift towards active transport.
Recognition of the importance of this shift for improved health and well-being
of citizens, and the wider challenge to decarbonise the global economy,
has grown with pressures on health and public transport systems during the pandemic.
The success of policies to accelerate cycling uptake depends on a range of
factors including, vitally, the design of the
cycling network for potential cyclists.
The approach presented in this paper provides a strong evidence base,
that considers both cycling potential and social equity,
for designing joined-up and cost-effective strategic cycle networks.
<!-- In tandem with policies to reduce excessive reliance on private cars, evidence-based investment in strategic cycle networks is key to. -->
